---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
library(ggplot2)
library(scales)
library(Hmisc)

# Loading and preprocessing the data

1. Load the activity data using read.csv
2. Transform the data if needed for analysis
```{r}
# Clear the work space
rm(list=ls())

# Load the raw activity data
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')

# format the date 
activity_raw$date <- as.POSIXct(activity_raw$date, format="%Y-%m-%d")

# Compute the weekdays from the date attribute
activity_raw <- data.frame(date=activity_raw$date, 
                           weekday=tolower(weekdays(activity_raw$date)), 
                           steps=activity_raw$steps, 
                           interval=activity_raw$interval)

# Compute the day type (weekend or weekday)
activity_raw <- cbind(activity_raw, 
                      daytype=ifelse(activity_raw$weekday == "saturday" | 
                                     activity_raw$weekday == "sunday", "weekend", 
                                     "weekday"))

# Create the final data.frame
activity <- data.frame(date=activity_raw$date, 
                       weekday=activity_raw$weekday, 
                       daytype=activity_raw$daytype, 
                       interval=activity_raw$interval,
                       steps=activity_raw$steps)

# Clear the workspace
rm(activity_raw) 
```

Display the first few rows of the activity data frame:
```{r}
head(activity)
```

# What is mean total number of steps taken per day?
1. Make a histogram of the total number of steps taken each day
2. Calculate and report the mean and median total number of steps taken per day
```{r}
# Compute the total number of steps each day (NA values removed)
sum_data <- aggregate(activity$steps, by=list(activity$date), FUN=sum, na.rm=TRUE)

# Rename the attributes
names(sum_data) <- c("date", "total")
```
Display the first few rows of the sum_data data frame:
```{r}
head(sum_data)
```
```{r}
# Compute the histogram of the total number of steps each day
hist(sum_data$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="green", 
     xlab="Total number of steps", 
     ylim=c(0, 20), 
     main="Histogram of the total number of steps taken each day\n(NA removed)")

# Calculate and report the mean and median total number of steps taken per day
mean(sum_data$total)
median(sum_data$total)
```

# What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
# Clear the workspace
rm(sum_data)

# Compute the means of steps accross all days for each interval
mean_data <- aggregate(activity$steps, 
                       by=list(activity$interval), 
                       FUN=mean, 
                       na.rm=TRUE)

# Rename the attributes
names(mean_data) <- c("interval", "mean")
```
Display the first few rows of the mean_data data frame:
```{r}
head(mean_data)
```
The time serie plot is created by the following lines of code
```{r}
# Compute the time series plot
plot(mean_data$interval, 
     mean_data$mean, 
     type="l", 
     col="green", 
     lwd=2, 
     xlab="Interval [minutes]", 
     ylab="Average number of steps", 
     main="Time-series of the average number of steps per intervals\n(NA removed)")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
# We find the position of the maximum mean
max_pos <- which(mean_data$mean == max(mean_data$mean))

# We lookup the value of interval at this position
max_interval <- mean_data[max_pos, 1]

# Clear the workspace
rm(max_pos, mean_data)
```
The 5-minute interval that contains the maximum of steps, on average across all days, is 
```{r}
max_interval
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
